Brief Announcement: An Improved Lower Bound for Dynamic Graph Partitioning

### 2020-01-22 paper/paper.tex
Minor rewrite of intro

Main technical contribution of this paper is $\Omega(k\ell)$ lower bound for any deterministic algorithm for Online Balanced Partition problem.
Previously the best known lower bound was $\Omega(k)$ \cite{repartition-disc}.
The gap between upper and lower bound is still substantial, as the best known algorithm is the component-based algorithm of   \cite{repartition-disc} with competitive ratio $O(k^2\ell^2)$.

We contribute to the search for the optimal algorithm by presenting $O(k\ell)$-competitive algorithm for the special case where the perfect partition of the input graph is assumed and every algorithm must colocate every communicating pair.
Note that our $\Omega(k\ell)$ lower bound uses a perfect partition input graph and colocates communicating pairs, and hence we provide a tight result for such scenario.

@@

In this paper we study an online balanced partitioning problem.
The static offline version of~the~problem, i.e., a problem variant where
migration is not allowed, where all requests are known in advance, and where
the goal is to find an assignment of $n$ nodes to $\ell$~physical machines, each of~capacity $n/\ell$, is known as the
\emph{$\ell$-balanced graph partitioning problem}. The problem is 
NP-complete, and cannot even be approximated within any finite factor unless P
= NP~\cite{AndRae06}.  The static
variant where $\ell = 2$ corresponds to the minimum bisection problem, which
is already NP-hard~\cite{GaJoSt76}, and 
the currently best approximation ratio is $O(\log n)$~\cite{SarVaz95,ArKaKa99,FeKrNi00,FeiKra02,KraFei06,Raec08}.
The inapproximability of the static variant for general values of $\ell$
motivated research on the bicriteria variant, which can be seen as the offline
counterpart of our capacity augmentation approach. Here, the~goal
is~to~compute a~graph partitioning into $\ell$ components of~size at most~$k$ (where $k > n/\ell$) and the cost of the cut is compared to the optimal (non-augmented)
solution where all components are of a~size at most $n/k$. The variant where
$n \geq 2 \cdot k \cdot \ell$ was considered in
\cite{LeMaTr90,SimTen97,EvNaRS00,EvNaRS99,KrNaSc09}. So far, the~best result~is~an~$O(\!\sqrt{\log n \cdot \log \ell})$-approximation algorithm~\cite{KrNaSc09}.

@@

h\item \emph{Where to migrate, and what?}
If an algorithm decides to collocate nodes $x$ and~$y$, the~question becomes
how. Should $x$ be migrated to the physical machine holding $y$, $y$ to the one holding
$x$, or should both nodes be migrated to a new machine?

@@

\section{Problem Definition}

@@

\section{Core results}
\subsection{Lower bound $\Omega(k\cdot \ell)$}

@@

\subsection{Perfect Partition Learner Algorithm}

@@

@@

>>>>

Main technical contribution of this paper is $\Omega(k\ell)$ lower bound for the competitive ratio of any deterministic algorithm for Online Balanced Partition problem.
The best known lower bound so far was $\Omega(k)$ \cite{repartition-disc}.

Additionally, we contribute to search for the optimal algorithm by presenting $O(k\ell)$-competitive algorithm for the special case, where the perfect partition of the input graph is assumed, and every algorithm must colocate every communicating pair.
Note that our $\Omega(k\ell)$ lower bound uses a perfect partition input graph and colocates communicating pairs, and hence for this scenario, our results are tight.
\maciek{Also, it is easy to construct the adversary that forces collocation of requested nodes}

@@

In this paper we study the Online Balanced Partitioning problem (formally presented in Section~\ref{sec:problem-definition}).
The static offline version of~the~problem, i.e., a problem variant where
migration is not allowed, where all requests are known in advance, and where
the goal is to find an assignment of $n$ nodes to $\ell$~physical machines, each of~capacity $n/\ell$, is known as the
\emph{$\ell$-balanced graph partitioning problem}. The problem is 
NP-complete, and cannot even be approximated within any finite factor unless P
= NP~\cite{AndRae06}.  The static
variant where $\ell = 2$ corresponds to the minimum bisection problem, which
is already NP-hard~\cite{GaJoSt76}, and 
the currently best approximation ratio is $O(\log n)$~\cite{SarVaz95,ArKaKa99,FeKrNi00,FeiKra02,KraFei06,Raec08}.
The inapproximability of the static variant for general values of $\ell$
motivated research on the bicriteria variant, which can be seen as the offline
counterpart of our capacity augmentation approach. Here, the~goal
is~to~compute a~graph partitioning into $\ell$ components of~size at most~$k$ (where $k > n/\ell$) and the cost of the cut is compared to the optimal (non-augmented)
solution where all components are of a~size at most $n/k$. The variant where
$n \geq 2 \cdot k \cdot \ell$ was considered in
\cite{LeMaTr90,SimTen97,EvNaRS00,EvNaRS99,KrNaSc09}. So far, the~best result~is~an~$O(\!\sqrt{\log n \cdot \log \ell})$-approximation algorithm~\cite{KrNaSc09}.

@@

\item \emph{Where to migrate, and what?}
If an algorithm decides to collocate nodes $x$ and~$y$, the~question becomes
how. Should $x$ be migrated to the physical machine holding $y$, $y$ to the one holding
$x$, or should both nodes be migrated to a new machine?

@@

\section{Problem Definition}
\label{sec:problem-definition}

@@

\section{Lower bound $\Omega(k\cdot \ell)$ for Competitive Ratio of Any Deterministic Algorithm}

@@

\section{Perfect Partition Learner Algorithm}

@@

\section{Corollary and Future Directions}

The gap between upper and lower bound is still substantial, as the best known algorithm is the component-based algorithm of   \cite{repartition-disc} with competitive ratio $O(k^2\ell^2)$.

\begin{appendix}

@@

\end{appendix}
